---
title: "Going through parts of the flowCore package to understand it"
author: "Thorsten Brach"
date: "2019/01/03"
output: html_document
---



#  Load all required packages Packages

- References: 
    - the flowCore vignette just google and find the newest
    - This course: [https://bioinformatics.ca/workshops/2013/flow-cytometry-data-analysis-using-r-2013]



```{r, echo = FALSE, message=FALSE, include = FALSE}
## flowCore:
biocLite("BiocUpgrade")
# source("https://bioconductor.org/biocLite.R")
# biocLite("flowCore")
library(flowCore); #packageVersion("flowCore")

## ggcyto: for the gating set functions and some plot options
# source("https://bioconductor.org/biocLite.R")
# biocLite("ggcyto")
library(ggcyto); #packageVersion("ggcyto")

## ggplot2
library(ggplot2); #packageVersion("ggplot2")

## dplyr
library(dplyr); #packageVersion("dplyr")

## lubridate
library(lubridate)

## gridExtra for grid.arrange
library(gridExtra)

```



# a flowFrame object

- The basic unit is a flowFrame corresponding to a single fcs file, it is a S4 object
    - the exprs slot contains the event data as a matrix
    - the parameters slot contains the meta data as an AnnotatedDataFrame with meta data derived from an FCS
file’s “$P¡n¿” keywords, which describe the detector and stain information

- as shown below, I understand how to get to the event matrix in the flowFrame object
- I wrote extract_parameter_keywords_from_flowFrame and extract_all_non_parameter_keywords_from_flowFrame that allow you to extract the meta data using the keyword function and present it in data frames

```{r}
# - define the path to the data you want to analyse -
datapath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/Data/20181129_TestExperiment_DTU"
# --
files <- list.files(datapath) 
files <- files[grepl("*.fcs", files)]

# - read the fcs file in -
file1 <- read.FCS(filename = file.path(datapath, files[2]), transformation=FALSE) # NB: alter.names = TRUE would change FSC-A to FSC.A so you would not always have to use backticks
# --

# - explore the fcs file -
isS4(file1) # it's an S4 object
# with @ you see the slots exprs, parameters, and description
head(file1@exprs) # shows you the events and what has been recorded
dim(exprs(file1)) # 200000 events and 11 variables
summary(file1) # quite nicely showing an overview of the variables

# -- look at the meta information --
names(file1) # basically just the column names
featureNames(file1) # bascically also the colnames
colnames(file1@exprs)
colnames(file1)
ADF <- file1@parameters
class(ADF)
dim(ADF) # so doesn't have that much informatoin
# the keyword method has I guess all information
KeywordList <- keyword(file1) # here 153 keywords
names(KeywordList)
# I wrote a little function to extract the keywords related to the parameters (the detectors)
ParameterValues <- extract_parameter_keywords_from_flowFrame(flowFrame = file1, suffixVector = NULL)
# and I also wrote a function to extract all non parameter keywords
NonParamterValues <- extract_all_non_parameter_keywords_from_flowFrame(file1)
# both functions produce a dataframe and can be extended hopefully with a flowSet object
# ----
```



# a flowSet object

- This class provides a mechanism for efficiently hosting the flowFrame objects with minimal copying, reducing memory requirements

```{r}
# - read several flowFrames into a list -
datapath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/Data/20181129_TestExperiment_DTU"
files <- list.files(datapath) 
files <- files[grepl("*.fcs", files)]

fcsFileList <- lapply(file.path(datapath, files[c(2,4,6)]), read.FCS, transformation=FALSE)
fcsFileList <- lapply(file.path(datapath, files), read.FCS, transformation=FALSE, column.pattern = "-W|-H", invert.pattern = TRUE) # NB: I needed to use the patterns here, since in some cases the -W and -H columns were not recorded, unfortunately it does not affect the P names, e.g. $P15 still exists in the ones with -H
names(fcsFileList) <- files
# --

# - Einschub: analyse the keyword parameters from the list to compare -
ParameterValuesList <- lapply(fcsFileList, extract_parameter_keywords_from_flowFrame) # if you look at this list, you see that these

# -- test that at least all but BS and MS are identical --
ParameterValuesList_Restricted <- lapply(ParameterValuesList, function(DF){
        DF <- DF[c("N", "R", "B", "E", "V", "G", "DISPLAY")]
        rownames(DF) <- NULL
        DF
})

length(unique(ParameterValuesList_Restricted)) # if 1 all experiments had used the same settings, then print only 1 otherwise all
# ----

nonParameterValuesList <- lapply(fcsFileList, extract_all_non_parameter_keywords_from_flowFrame)

# -- check that the length are equal (better would be to check that all Keywords are identical) --
if(length(unique(vapply(nonParameterValuesList, nrow, numeric(1)))) == 1) {
        nPV_DF <- do.call("cbind", nonParameterValuesList)
        nPV_DF <- nPV_DF[, c(1, seq(2, ncol(nPV_DF), by = 2))]
        
} else {
        nPV_DF <- NULL
}
# ----
# --

# - got to the flowSet object -
fs <- as(fcsFileList, "flowSet")

# fs has additional meta info that was not in the flowFrames
fs@phenoData
# you can add info here and learn the fsApply method:
phenoData(fs)$Filename <- fsApply(fs,keyword, "$FIL")
fs@phenoData
class(fs@phenoData) # again an AnnotatedDataFrame
# some convenience functions
fsApply(fs, each_col, median)
fsApply(fs, each_col, min)

```

- So clearly there is more info in the vignette, but I skip this now a bit to get to the parts that I might actually use



# Gating


- Like transformations, flowCore includes a number of built-in common flow cytometry gates. The
simplest of these gates are the geometric gates
- In addition, we introduce the notion of data-driven gates, or filters, not usually found in flow cytometry
software. In these approaches, the necessary parameters are computed based on the properties
of the underlying data, for instance by modelling data distribution or by density estimation
- but I also skip this section because in the end it seems it just becomes the gating set option


# Gating Set

- for next time

```{r}
```