---
title: "Template evaluation of flow cytometry data for cell counting project"
author: "Thorsten Brach"
date: "2019/01/03"
output: html_document
---




# Possible improvements and own reminders

- analyse counting beads with regard to average flow per second

- FC.geom_hex with a high enough bin is just as good as FC.geom_point (no legend) and FC.geom_point_kernel. So I recommend geom_hex, it is easy and fast, but the point functions work fine two, I recommend kernel version but it is a bit slow.
    - NB: lines in the density plot make sense, since the data consists of discrete steps!




# To use this template, you have to:

- for a new experiment just adjust the section USER INPUT: Set parameters
- All Thresholds at the LSRII refer to width, e.g. FSC-W and SSC-W and APC-W, it seems at DTU it is H (FSC-H was min 497 and SSC-H was 193 for THRESHOLDS 500 and 200 respectively)
- Area = Width*Height/(2^16) at least for FSC-A and SSC-A, but most likely also for PE-A and FITC-A, because they take the same "steps"




# Packages

- References: 
    - the flowCore vignette [https://www.bioconductor.org/packages/devel/bioc/vignettes/flowCore/inst/doc/HowTo-flowCore.pdf]
    - This course: [https://bioinformatics.ca/workshops/2013/flow-cytometry-data-analysis-using-r-2013]


```{r, echo = FALSE, message=FALSE, include = FALSE}
## MASS is for the 2D kernel function for density coloring <https://slowkow.com/notes/ggplot2-color-by-density/>
# library(MASS)

## flowCore:
# biocLite("BiocUpgrade")
# source("https://bioconductor.org/biocLite.R")
# biocLite("flowCore")
library(flowCore); #packageVersion("flowCore")

## ggcyto: for the gating set functions and some plot options
# source("https://bioconductor.org/biocLite.R")
# biocLite("ggcyto")
library(ggcyto); #packageVersion("ggcyto")

## ggplot2
library(ggplot2); #packageVersion("ggplot2")

## dplyr
library(dplyr); #packageVersion("dplyr")

## dplyr
library(tidyr); #packageVersion("tidyr")

## lubridate
library(lubridate)

## gridExtra for grid.arrange
library(gridExtra)

## viridis is just for colors
library(viridis)
```




# USER INPUT: Set parameters


```{r, echo = TRUE, message=FALSE}
# - define the path to your function file -
functionpath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/Functions_FC"
function_file <- "20190103_FlowCytometry_Functions.R"
# --


# - define the path to the data you want to analyse -
datapath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/Data/20190104_DTUExample"
# --


# - gating Thresholds and Polygons -
# -- for the area gate --
FSC.A.low <- 0.5
FSC.A.high <- Inf
SSC.A.low <- 0.5
SSC.A.high <- Inf
# ----

# -- for gating on Pacific Blue-A FSC-A --
Poly <- data.frame(x = c(200, 15000, 15000, 2000, 400), y = c(200, 200, 7000, 7000, 800))
# ----

# -- for gating on PE-A FSC-A for the beads --
Poly_Beads <- data.frame(x = c(20000, 260000, 260000, 20000), y = c(30000, 30000, 260000, 260000))
# ----
# --


# - plotting parameters -
bins <- 300 # defines how "fine" the FC.geom_hex plots will be, the higher the number of bins the more the plot looks like a geom_point plot (a la flowjo) but it's still faster
# --


# - IMPORTANT: Check carefully: Protocol and flow cytometer parameters used -
# -- protocol volumes and concentrations
muL_Buffer_per_mg_feces <- 10 # e.g. 10 when you add 1 ml to 100 mg feces
muL_supernatant <- 200 # the amount of supernatant used after the slow spin, NB: not needed for calculation but record it anyway
muL_Transfer1 <- 20
muL_Buffer_Transfer1 <- 2980
muL_Transfer2 <- 440
muL_Stain <- 50 # used in Transfer 1
beads_concentration_permuL <- 1022 # in the bought bead tube
beads_volume_muL <- 10 # added in Transfer 2
# ----

# -- flow cytometer settings --
set_flow_rate_muL_s <- 0.5
# ----
# --


# - inputs for saving the data -
savepath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/Results"
savename <- NULL
# --
```




# Source the functions


```{r, echo = FALSE, message=FALSE}
source(file.path(functionpath, function_file))
```




# Load the fcs files as flowFrame object list


```{r, echo = FALSE, message=FALSE}
files <- list.files(datapath) 
files <- files[grepl("*.fcs", files)]

# fcsFileList <- lapply(file.path(datapath, files), read.FCS, transformation=FALSE)
fcsFileList <- lapply(file.path(datapath, files), read.FCS, transformation=FALSE, column.pattern = "-W|-H", invert.pattern = TRUE) # NB: I needed to use the patterns here, since in some cases the -W and -H columns were not recorded, unfortunately it does not affect the P names, e.g. $P15 still exists in the ones with -H
names(fcsFileList) <- files
```




# Analyse the keyword parameters of the loaded files



## The parameters related to the data channels

```{r, echo = FALSE, message=FALSE}
ParameterValuesList <- lapply(fcsFileList, extract_parameter_keywords_from_flowFrame) 

# -- test that at least all but BS and MS are identical --
ParameterValuesList_Restricted <- lapply(ParameterValuesList, function(DF){
        DF <- DF[c("N", "R", "B", "E", "V", "G", "DISPLAY")]
        rownames(DF) <- NULL
        DF
})

if (length(unique(ParameterValuesList_Restricted))) {
        
        PrintList <- ParameterValuesList[[1]]
} else {
        PrintList <- ParameterValuesList
}
```

- **The parameter based keywords are summarized in the following table/tables**

```{r, echo = FALSE, message=FALSE}
PrintList
```



## The parameters not directly related to the data channels

```{r, echo = FALSE, message=FALSE}
nonParameterValuesList <- lapply(fcsFileList, extract_all_non_parameter_keywords_from_flowFrame)

# -- check that the length are equal (better would be to check that all Keywords are identical) --
if(length(unique(vapply(nonParameterValuesList, nrow, numeric(1)))) == 1) {
        nPV_DF <- do.call("cbind", nonParameterValuesList)
        nPV_DF <- nPV_DF[, c(1, seq(2, ncol(nPV_DF), by = 2))]
        
} else {
        nPV_DF <- NULL
}
```

```{r, echo = FALSE, message=FALSE}
knitr::kable(nPV_DF, caption = "Table of non-parameter related keywords")
```




# Change the data into a flowSet object and extract more compact keyword info


```{r, echo = FALSE, message=FALSE}
fs <- as(fcsFileList, "flowSet")

KeywordTable <- as.data.frame(keyword(fs, c("$TOT", "$PAR", "TUBE NAME", "THRESHOLD", "$DATE", "EXPORT TIME", "$BTIM", "$ETIM")))

# change sample Names
# flowWorkspace::sampleNames(fs) <- paste(Prefix, fs@phenoData@data$well, fs@phenoData@data$SRC, sep = "_")
# 
# rownames(KeywordTable) <- flowWorkspace::sampleNames(fs)

KeywordTable$`$BTIM` <- hms(KeywordTable$`$BTIM`)
KeywordTable$`$ETIM` <- hms(KeywordTable$`$ETIM`)
KeywordTable$RecTime_sec <- as.numeric(as.duration(KeywordTable$`$ETIM`- KeywordTable$`$BTIM`), "seconds")
# - new add the run time from the exprs data -
Maximums <- as.data.frame(fsApply(fs, each_col, max))
KeywordTable$RunTime_sec <- Maximums$Time/100
# --
```


```{r, echo = FALSE, message=FALSE}
knitr::kable(KeywordTable, caption = "summary of key keywords with running times added.")
```




# Look at the Minimal values and steps (above 0) to understand thresholding and to possibly adjust area gates

- shows you usually that thresholding is not on area but width or height


```{r, echo = FALSE, message=FALSE}
# Determine Minimum values of the different parameters in the different samples
MatrixList <- lapply(1:length(fs), FUN = function(x){exprs(fs[[x]])})
names(MatrixList) <- flowWorkspace::sampleNames(fs)
Minimums <- t(sapply(MatrixList, function(x){apply(x, 2, min)}))
# Minimums <- fsApply(fs, each_col, min)
MinimumsAbove0 <- t(sapply(MatrixList, function(x){apply(x, 2, function(y) min(y[y>0]))}))
# MinimumsAbove0 <- fsApply(fs, each_col, function(x){min(x[x>0])})


BigDF <- as.data.frame(do.call("rbind", MatrixList))

Discretized <- lapply(BigDF, function(x){head(sort(unique(x[x>0])), 10)})

Discretized <- do.call("cbind", Discretized)

rm(BigDF)
```


```{r, echo = FALSE, message=FALSE}
# Determine Minimum values of the different parameters in the different samples
knitr::kable(Minimums, caption = "Minimums")
knitr::kable(MinimumsAbove0, caption = "Minimums above 0")
knitr::kable(Maximums, caption = "Maximums")
knitr::kable(Discretized, caption = "Discretized steps above 0")
```




# Gating of cells and beads



## Transform flowSet object to GatingSet object

```{r, echo = FALSE, message=FALSE}
gs <- GatingSet(fs) # also GatingSets are S4 objects
```



## Gate 1: Area: Remove Events with negative FSC-A or SSC-A, i.e events that stay are: FSC-A > `r FSC.A.low` and < `r FSC.A.high`, SSC-A > `r SSC.A.low` and < `r SSC.A.high`.  

- NB: Events with negative areas have either width or height = 0. 
- In rare cases Area is however not < 0 when width is 0 but height is max (specifically SSC-W could be 0 for SSC-A not below 0). I would therefore add here FSC-H/W SSC-H/W 1 Inf gates, but still call it just area gate (**NB: later you might gate on higher FSC-H min values, the W gate here is silly since you use higher thresholds**). Remember, for areas above 0 the equation is area = width*height/2^16


```{r, message = FALSE, warning = FALSE}
Gate.Area <- rectangleGate(filterId = "Area", list("FSC-A" = c(FSC.A.low, FSC.A.high), "SSC-A" = c(SSC.A.low, SSC.A.high)))
# add the gate to the gs
add(gs, Gate.Area, parent = "root")
# gate the data
recompute(gs)

gatingStats <- gs.Stats(gs)
```


```{r, echo = FALSE, message=FALSE}
knitr::kable(gatingStats[[1]], caption = "after gate 1")
knitr::kable(gatingStats[[2]], caption = "after gate 1")
```




## Gate 2: Polynomial gate on Pacific Blue-A FSC-A plots



### Pacific Blue-A FSC-A plots with polygon

- Pacific Blue-A detects the DAPI


```{r, echo = FALSE, message=FALSE}
TrL <- FC.geom_hex(fs = getData(gs, "/Area"), x = "Pacific Blue-A", y = "FSC-A", bins = bins, co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", colPalette = NULL, ncolors = 10)

# - geom_point alternatives, are slower and with high enough bins FC.geom_hex contains almost all info -
# TrL_p <- FC.geom_point(fs = getData(gs, "/Area"), x = "Pacific Blue-A", y = "FSC-A", pch = ".", co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", nbin = 150)
# # 
# # # -- even slower but with legend, so I prefer this point alternative --
# TrL_p <- FC.geom_point_kernel(fs = getData(gs, "/Area"), x = "Pacific Blue-A", y = "FSC-A", pch = ".", co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", colPalette = viridis, nbin = 70, ncolors = 10)
# # # ----
# --

#Poly <- data.frame(x = c(200, 15000, 15000, 2000, 400), y = c(200, 200, 7000, 7000, 800))
TrL_Poly <- plotPolygon(TrL = TrL, polyDF = Poly)
```


```{r, fig.width= 14, fig.height = 3.75*length(gs)/2, warning = FALSE, message = FALSE}
do.call("grid.arrange", c(TrL_Poly, ncol = 2))
```



### Apply gate 2


```{r, warning = FALSE, message = FALSE}
matr <- as.matrix(Poly)
colnames(matr) <- c("Pacific Blue-A", "FSC-A")
Gate.Poly <- polygonGate(matr, filterId = "Cells")
# add the gate to the gs
add(gs, Gate.Poly, parent = "/Area")
# gate the data
recompute(gs)

gatingStats <- gs.Stats(gs)
```


```{r, echo = FALSE, message=FALSE}
knitr::kable(gatingStats[[1]], caption = "after gate 2")
knitr::kable(gatingStats[[2]], caption = "after gate 2")
```




### Density FSC-A SSC-A plots after the poly gate

- **just showed that you could not get rid of the remaining buffer only values (but well with the used threshold there were anyway not many buffer only events)**

```{r, echo = FALSE, message=FALSE}
# TrL <- FC.geom_hex(fs = getData(gs, "/Area/Poly"), x = "FSC-A", y = "SSC-A", bins = 40, co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10")

```


```{r, fig.width= 14, fig.height = 3.75*length(gs)/2, warning = FALSE, message = FALSE}
# do.call("grid.arrange", c(TrL, ncol = 2))
```



## Gate 3: Polynomial/Rectangular gate on PE-A FSC-A plots to detect the beads



### PE-A FSC-A plots with polygon

- PE-A detects the beads


```{r, echo = FALSE, message=FALSE}

TrL <- FC.geom_hex(fs = getData(gs, "/Area"), x = "PE-A", y = "FSC-A", bins = bins, co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", colPalette = NULL, ncolors = 10)

# - geom_point alternatives, are slower and with high enough bins FC.geom_hex contains almost all info -
# TrL_p <- FC.geom_point(fs = getData(gs, "/Area"), x = "PE-A", y = "FSC-A", pch = ".", co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", nbin = 150)
# 
# # -- even slower but with legend, so I prefer this point alternative --
# TrL_p <- FC.geom_point_kernel(fs = getData(gs, "/Area"), x = "PE-A", y = "FSC-A", pch = ".", co.fix = TRUE, xlimits = c(0.1, 300000), ylimits = c(.1,300000), trans = "log10", colPalette = viridis, nbin = 70, ncolors = 10)
# # ----
# --


TrL_Poly <- plotPolygon(TrL = TrL, polyDF = Poly_Beads)
```


```{r, fig.width= 14, fig.height = 3.75*length(gs)/2, warning = FALSE, message = FALSE}
do.call("grid.arrange", c(TrL_Poly, ncol = 2))
```



### Apply gate 3


```{r, echo = FALSE, warning = FALSE, message = FALSE}
matr <- as.matrix(Poly_Beads)
colnames(matr) <- c("PE-A", "FSC-A")
Gate.Poly <- polygonGate(matr, filterId = "Beads")
# add the gate to the gs
add(gs, Gate.Poly, parent = "/Area")
# gate the data
recompute(gs)

gatingStats <- gs.Stats(gs)
```


```{r, echo = FALSE, message=FALSE}
knitr::kable(gatingStats[[1]], caption = "after gate 3")
knitr::kable(gatingStats[[2]], caption = "after gate 3")
```



## Gate Illustration

```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot(gs)
```




# Illustrate the flow and the arrival of the beads


- simple plots illustrating how the number of events changed over time. If the flow was smooth one would expect these curves to be smooth


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# remember fs is still the completely original flowSet, i.e. fs <- getData(gs, "/root")
fs_beads <- getData(gs, "/Area/Beads")
MatrixList_Beads <- lapply(1:length(fs_beads), FUN = function(x){exprs(fs_beads[[x]])})
names(MatrixList_Beads) <- flowWorkspace::sampleNames(fs_beads)

# - change Matrixes to data frames -
DF_List <- lapply(MatrixList, as.data.frame)
DF_List_Beads <- lapply(MatrixList_Beads, as.data.frame)
# --


# - Goal 1: Event plots where the bead events are marked - 
# for this you have to find the bead events in the total events, here is how it can be done:

# -- put bead DF on top of event DF -
DFF_List <- Map(f = rbind, DF_List_Beads, DF_List)
# ----

# -- find duplicated events in combined DFFs --
# takes a bit of time but not brutally long
DFF_List <- lapply(DFF_List, function(DFF){
        DFF$Duplicated <- duplicated(DFF)
        DFF
})
# ----

# -- remove the bead DFs again --
no_Beads <- sapply(DF_List_Beads, nrow)
DFF_List <- lapply(1:length(DFF_List), function(i){
        DFF <- DFF_List[[i]]
        DFF <- DFF[(no_Beads[i]+1):nrow(DFF), ]
        DFF$Type <- "event"
        DFF$Type[DFF$Duplicated] <- "bead"
        DFF
})
if (! identical(names(DF_List), names(DF_List_Beads))) {stop("beads and event DF did not fit together")}
names(DFF_List) <- names(DF_List)
#all.equal(sapply(DFF_List, function(DFF){sum(DFF$Duplicated)}), no_Beads)
if(!identical(sapply(DFF_List, nrow), sapply(DF_List, nrow))) {stop("you changed the number of rows here")}
# ----

# -- count the events and change the time to seconds --
DFF_List <- lapply(DFF_List, function(DFF){
        DFF$Events <- 1:nrow(DFF)
        DFF$Time <- DFF$Time/100
        DFF
})
# ----


# -- Do linear fits and record the R2 to add later to the plots --
R_squares_all_events <- lapply(DFF_List, function(DFF){
        fit <- lm(Events ~ Time, data = DFF)
        summary(fit)$adj.r.square
})
# --


# -- define plot functions --
plot_flow <- function(DF) {
        DF2 <- dplyr::filter(DF, Type == "bead")
        Tr <- ggplot(DF, aes(x = Time, y = Events))
        Tr <- Tr +
                geom_point(data = DF2, size = 2, col = "red") +
                geom_smooth(method = "lm", se = FALSE) +
                geom_line() +
                xlab("Time [s]") # +
                # theme_bw()
        Tr
}
# ----


# -- plot the events --
TrL <- lapply(DFF_List, plot_flow)

TrL <- lapply(1:length(TrL), function(i) {
        Tr <- TrL[[i]]
        Tr <- Tr +
                ggtitle(paste0(names(DFF_List)[i], ": r2: ", round(R_squares_all_events[[i]], 4)))
        Tr})
# ----
# --



# - Goal 2: Plot only the arrival of the beads - 
# making sure you can deal with samples without beads
no_Beads <- sapply(DF_List_Beads, nrow)

DF_List_Beads <- lapply(MatrixList_Beads, as.data.frame)


# -- count the events and change the time to seconds --
DF_List_Beads <- lapply(DF_List_Beads, function(DFF){
        if (nrow(DFF) > 0) {
                DFF$Events <- 1:nrow(DFF)
                DFF$Time <- DFF$Time/100
                DFF
        } else {
                DFF$Events <- numeric(0)
                DFF
        }
        
})
# ----


# -- Do linear fits and record the R2 and the slopes to add later to the plots --
bead_r2_slopes <- lapply(DF_List_Beads, function(DFF){
        if (nrow(DFF) > 0) {
                fit <- lm(Events ~ Time, data = DFF)
                r2 <- summary(fit)$adj.r.square
                slope <- fit$coef[2]
                data.frame(r2 = r2, slope = slope)
        } else {
                r2 <- 0
                slope <- 0
                data.frame(r2 = r2, slope = slope)
        }
        
})

bead_r2_slopes <- do.call("rbind", bead_r2_slopes)
# --


# -- define plot functions --
plot_flow_beads <- function(DF) {
        Tr <- ggplot(DF, aes(x = Time, y = Events))
        Tr <- Tr +
                geom_point(size = 2, col = "red") +
                geom_smooth(method = "lm", se = FALSE) +
                geom_line() +
                xlab("Time [s]") # +
                ylab("Beads")
                # theme_bw()
        Tr
}
# ----

# -- plot the events --
TrL_Beads <- lapply(DF_List_Beads, plot_flow_beads)

TrL_Beads <- lapply(1:length(TrL_Beads), function(i) {
        Tr <- TrL_Beads[[i]]
        Tr <- Tr +
                ggtitle(paste0(names(DF_List_Beads)[i], "; r2: ", round(bead_r2_slopes$r2[i], 4), ", slope: ", round(bead_r2_slopes$slope[i], 3)))
        Tr})
# ----
# --


# - combine the two lists so they alternate -
TrLL <- c(rbind(TrL, TrL_Beads))
# --
```



```{r, fig.width= 14, fig.height = 3.75*length(TrLL)/2, warning = FALSE, message = FALSE}
do.call("grid.arrange", c(TrLL, ncol = 2))
```




# Generate a Result DF

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - prepare a result DF -
CountFreq <- gs.Stats(gs, gates = getNodes(gs))
res_DF <- cbind(KeywordTable, CountFreq[[1]])
res_DF <- cbind(Sample = rownames(res_DF), res_DF)


# -- add all the protocol inputs NB: Here the code strictly depends on your input--
res_DF <- dplyr::mutate(res_DF, muL_Buffer_per_mg_feces = muL_Buffer_per_mg_feces, muL_Transfer1 = muL_Transfer1,
                        muL_Buffer_Transfer1 = muL_Buffer_Transfer1, muL_Transfer2 = muL_Transfer2,
                        muL_Stain = muL_Stain, beads_conc_permuL = beads_concentration_permuL, 
                        beads_vol_muL = beads_volume_muL, sample_vol_muL = muL_Transfer2 + muL_Stain + beads_volume_muL)
# ----

# -- calculate beads_per_muL in the flow --
res_DF <- dplyr::mutate(res_DF, beads_per_muL_flow = beads_conc_permuL*beads_vol_muL/sample_vol_muL)
# ----
# -- calculate run_volume_muL and Cells_per_muL_flow --
res_DF <- dplyr::mutate(res_DF, run_volume_muL = Beads/beads_per_muL_flow, Cells_per_muL_flow = Cells/run_volume_muL)
# ----
# -- Calculate dilution factors --

res_DF <- dplyr::mutate(res_DF, DilFac_Transfer2 = (muL_Transfer2 + muL_Stain + beads_vol_muL)/muL_Transfer2,
                        DilFac_Transfer1 = (muL_Transfer1 + muL_Buffer_Transfer1)/muL_Transfer1,
                        Cells_per_muL_undiluted = Cells_per_muL_flow * DilFac_Transfer1 * DilFac_Transfer2,
                        Cells_per_mg_feces = Cells_per_muL_undiluted * muL_Buffer_per_mg_feces,
                        Cells_per_g_feces = Cells_per_mg_feces * 1000)   
# ----
# --
```



## Check the consistency of the flow rates derived from the beads, add to result DF (Alternative volume determination)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - add flow rates to result DF -
res_DF <- dplyr::mutate(res_DF, flow_muL_per_sec_from_beads = run_volume_muL/RunTime_sec, 
                        bead_slopes = bead_r2_slopes$slope,
                        flow_muL_per_sec_from_bead_slopes = bead_slopes/beads_per_muL_flow)  
# --

# - plot the flow rates -
plotDF <- dplyr::select(res_DF, Sample, flow_muL_per_sec_from_beads, flow_muL_per_sec_from_bead_slopes)
mean_slopes <- mean(plotDF$flow_muL_per_sec_from_bead_slopes)
mean_counts <- mean(plotDF$flow_muL_per_sec_from_beads)
plotDF <- tidyr::gather(plotDF, key = "type", value = "value", -Sample)
plotDF$type <- factor(plotDF$type, levels <- unique(plotDF$type), ordered = T)
col_values <- viridis(2)
names(col_values) <- levels(plotDF$type)
Tr <- ggplot(plotDF, aes(x = Sample, y = value, col = type))
Tr <- Tr +
        geom_hline(yintercept = set_flow_rate_muL_s) +
        geom_hline(yintercept = mean_counts, col = col_values[1], lty = 4) +
        geom_hline(yintercept = mean_slopes, col = col_values[2], lty = 4) +
        geom_point() +
        scale_color_manual("", values = col_values) +
        # facet_wrap(. ~ type) +
        xlab("") +
        ylab("flow rate [muL per s]") +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# --

```

- The following plot shows the flow rates calculated from the beads, the black horizontal line shows the flow rate that was set at the machine (user input)


```{r, fig.width= 10, fig.height = 6, warning = FALSE, message = FALSE}
Tr
```




# Generate a Result List for saving

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - prepare the result list to save all important data -
InputList <- list(functionpath = functionpath, function_file = function_file, datapath = datapath, 
                  FSC.A.low = FSC.A.low, FSC.A.high = FSC.A.high, SSC.A.low = SSC.A.low,
                  SSC.A.high = SSC.A.high, Poly = Poly, Poly_Beads = Poly_Beads,
                  bins = bins, muL_Buffer_per_mg_feces = muL_Buffer_per_mg_feces,
                  muL_supernatant = muL_supernatant, muL_Transfer1 = muL_Transfer1,
                  muL_Buffer_Transfer1 = muL_Buffer_Transfer1,
                  muL_Transfer2 = muL_Transfer2, 
                  muL_Stain = muL_Stain,
                  beads_concentration_permuL = beads_concentration_permuL,
                  beads_volume_muL = beads_volume_muL,
                  set_flow_rate_muL_s = set_flow_rate_muL_s,
                  savepath = savepath, savename = savename)


ValueDistributionList <- list(Minimums = Minimums, MinimumsAbove0 = MinimumsAbove0, Maximums = Maximums, Discretized = Discretized)

resList <- list(InputList = InputList, res_DF = res_DF, CountFreq = CountFreq, ParameterValuesList = ParameterValuesList, nonParameterValuesList = nonParameterValuesList, KeywordTable = KeywordTable)
# --
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::kable(res_DF, digits = 2, caption = "the final result data frame")
```




# Save the result list

```{r, echo = FALSE, message = FALSE, warning = FALSE}
if (!is.null(savepath)){
        
        if (is.null(savename)) {
                
                savename <- paste0(basename(datapath), "_results.rds")
        }
        
        saveRDS(object = resList, file.path(savepath, savename))
        
}
```










