# Count analysis person 1



## Background:

- The raw fcs files are saved in PILOT_Data folder
- The raw fcs files were analysed using the Rmd template: 20190305_FC_Analysis_LSR_PilotStudy_NewProtocol_Template.Rmd
    - The actually used Rmd files and the produced HTMLs can be found in the folder PILOT_Rmds_HTMLs
    - The runs resulted in actual count data that was saved as rds files in PILOT_result_rds_files
- Here we start the analysis with these rds files
    - We also use the excel spread sheets of the 5 participants where they noted down when they took the sample and what was the Bristol Stool score and so on. This data is in the folder PILOT_SampleCollection_SpreadsheetTemplates. I exported the information as csv files that is under Input_Data > SampleCollection_csvs
    


## Load functions and set colors
<!-- all lines where user user input is required or possible are marked with NB: USER INPUT and NB: user input, respectively -->


```{r, echo = FALSE, message=FALSE}
# - source all functions -
functionpath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/Functions_FC/"
function_file <- "20190103_FlowCytometry_Functions.R"
source(file.path(functionpath, function_file))
# --

# - for ggpubr -
symnum.args = list(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))
# -- 

# - Set Colors -
aliquot_colors <- QuantColors15[1:3]
names(aliquot_colors) <- c("Aliquot 1", "Aliquot 2", "Aliquot 3")

experimentor_shapes <- c(16, 17)
names(experimentor_shapes) <- c("Liwei", "Sandra")

# person_colors <- c(cbPalette[2:4], cbPalette[7:8])
# names(person_colors) <- levels(DF$Person)
thaw_colors <- c("#d3d3d3", cbPalette[1])
names(thaw_colors) <- c("FT", "ST")
# --
```



## Load the count rds files

```{r, echo = FALSE, message=FALSE, warning=FALSE}
datapath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/PILOT_result_rds_files/"
files <- list.files(datapath)
sampleName <- "20190213_PilotStudy_Person1_results.rds"


resL <- readRDS(file = file.path(datapath, sampleName))
res_DF <- resL[["res_DF"]]
```



## Define a Look Up table to assign the Wells of the run to Person, Timepoint, Aliquot and so on. Combine count data with the LookUp

- NB: The general plate layout can be found in the excel 20190131_Sample_and_Plate_Distribution_PilotStudy. There might be small changes in the layout for one or the other person, but in general this was talked through with Sandra and Liwei and checked. 


```{r, echo = FALSE, message = FALSE, warning = FALSE}
Person <- "P1"
Experiment_Date <- "2019-02-13"

LookUp <- data.frame(Tube = c(paste0("A", 1:5),
                              paste0("A", 11:12),
                              paste0("B", 1:12),
                              paste0("C", 1:12),
                              paste0("D", 1:12),
                              paste0("E", 1:12), 
                              paste0("F", 1:12), 
                              paste0("G", 1:12),
                              paste0("H", 1:3),
                              paste0("H", 9:11)), 
                     Person = c(rep("Control", 2), 
                                rep(Person, 2),
                                rep("Control", 1),
                                rep(Person, 74),
                                rep("Control", 2),
                                "Buffer",
                                rep(Person, 3)),
                     Timepoint = c(NA, NA, "t1", "t4", NA, "t3", "t3",
                                   rep(c(rep("t1", 3), rep("t2", 3), rep("t1", 3), rep("t2", 3)), 3),
                                   rep(c(rep("t3", 3), rep("t4", 3), rep("t3", 3), rep("t4", 3)), 3),
                                   NA, NA, NA, "t3", "t3", "t3"),
                     Thaw = c(NA, NA, "FT", "ST", NA, "FT", "FT",
                              rep(c("FT", "FT", "ST", "FT", "FT", "ST", "FT", "ST", "ST", "FT", "ST", "ST"), 6), 
                              NA, NA, NA, "FT", "FT", "FT"),
                     Aliquot = c("Aliquot 1", "Aliquot 2", "Aliquot 1", "Aliquot 3", "Aliquot 1", "Aliquot 2", "Aliquot 2",
                               rep(c(rep(c("Aliquot 1", "Aliquot 3", "Aliquot 2"), 2), rep(c("Aliquot 2", "Aliquot 1", "Aliquot 3"), 2)), 6),
                               "Aliquot 1", "Aliquot 2", NA, "Aliquot 2", "Aliquot 2", "Aliquot 2"),
                     Replicate = c(NA, NA, NA, NA, "r1", NA, NA,
                                   rep("r1", 12),
                                   rep("r2", 12),
                                   rep("r3", 12),
                                   rep("r1", 12),
                                   rep("r2", 12),
                                   rep("r3", 12),
                                   "r2", "r2", NA, "r1", "r2", "r3"),
                     Stain = c(rep("unstained", 4), "stained", "unstained", "unstained", rep("stained", 78)),
                     Experimentor = c(rep(c("Liwei", "Sandra"), 3), "Sandra", 
                                      rep(c(rep("Liwei", 6), rep("Sandra", 6)), 6),
                                      "Liwei", "Sandra", "Liwei", "Sandra", "Sandra", "Sandra"),
                     ExperimentDate = Experiment_Date)


# - Combine res_DF with LookUp -
colnames(res_DF)[5] <- "Tube"
DF <- merge(res_DF, LookUp, by = "Tube")
# --
```



## Load the sample collection data (Date, Bristol Score) and add to DF


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - Load Date and Time of time points and add that information -
sample_collection_path <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/PILOT_AfterFlow_Analysis_FInalCounts/Input_Data/SampleCollection_csvs/"
stool_data <- read.csv2(file = file.path(sample_collection_path, "00_Template_SampleCollectionSheet_P1.csv"), header = TRUE, sep = ";",
                        stringsAsFactors = FALSE)
stool_data$Date <- lubridate::parse_date_time(stool_data$Date, orders = "dmy", tz = "CET")
stool_data$Time <- lubridate::parse_date_time(stool_data$Time, orders = "HM", tz = "CET")
stool_data$Date_Time <-  update(stool_data$Date, hour = hour(stool_data$Time), minute = minute(stool_data$Time))

#colnames(stool_data) %in% colnames(DF)

DF <- merge(DF, stool_data, by = "Timepoint", all = TRUE)
# --
```



## Plot the counts (one where the sample volume was judged from beads, and one where it was judged from the set flow rate (preferred))

- I prefer the set flow rate because it appeared more stable, look at the HTMLs or go through 20190305_FC_Analysis_LSR_PilotStudy_NewProtocol_Template.Rmd
- I therefore also only show the plot of the counts from the flow rate



```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - set factors -
DF$Experimentor <- factor(DF$Experimentor, c("Liwei", "Sandra"), ordered = T)
DF$Aliquot <- factor(DF$Aliquot, c("Aliquot 1", "Aliquot 2", "Aliquot 3"), ordered = T)
# --

# - plot the counts using beads as volume -
DF_Plot <- DF[grepl(pattern = "B|C|D|E|F|G", DF$Tube), ] # so exclude all control and buffer counts in these plots


# First plot the number from the beads that might go up for t3 and t3
Tr <- ggplot(DF_Plot, aes(x = Timepoint, y = Cells_per_g_feces))

Tr <- Tr +
        geom_boxplot(aes(fill = Thaw), outlier.color = NA) +
        geom_point(aes(fill = Thaw, col = Aliquot, shape = Experimentor), position = position_jitterdodge(jitter.height = 0, jitter.width = 0.05), size = 3, alpha = 0.9) +
        theme_bw() +
        scale_color_manual("", values = aliquot_colors) +
        scale_fill_manual("", values = thaw_colors) +
        scale_shape_manual("", values = experimentor_shapes) +
        scale_y_continuous(limits = c(0, max(DF_Plot$Cells_per_g_feces))) +
        xlab("") +
        ylab("Cells [per g feces] (Volume from beads)") 


# pdf(file = "TimePoint_Plot_VolumeFromBeads_Person1.pdf", width = 6, height = 4.5)
# Tr
# dev.off()
# --


# - plot the counts using flow rate determined volume -
Tr <- ggplot(DF_Plot, aes(x = Timepoint, y = Cells_per_g_feces_FR))

Tr <- Tr +
        geom_boxplot(aes(fill = Thaw), outlier.color = NA) +
        geom_point(aes(fill = Thaw, col = Aliquot, shape = Experimentor), position = position_jitterdodge(jitter.height = 0, jitter.width = 0.05), size = 3, alpha = 0.9) +
        theme_bw() +
        scale_color_manual("", values = aliquot_colors) +
        scale_fill_manual("", values = thaw_colors) +
        scale_shape_manual("", values = experimentor_shapes) +
        scale_y_continuous(limits = c(0, max(DF_Plot$Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces] (Volume from set flow rate)") 


# pdf(file = "TimePoint_Plot_VolumeFromFR.pdf", width = 6, height = 4.5)
# Tr
# dev.off()
# --
```


```{r, echo = FALSE, message = FALSE, fig.width = 8, figh.height = 7, warning = FALSE}
Tr
```


## For Person 1 we had the extra test glass vs plastic

- Because our numbers were so much lower than the ones from Belgien, we wanted to test wether many bacteria might get stuck on the plastic


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - plot today the comparison plastic to glass -
DF_Plot2 <- DF[DF$Tube %in% c("E7", "F7", "G7", "H9", "H10", "H11"), ]
DF_Plot2$Material <- "Glass"
DF_Plot2$Material[1:3] <- "Plastic"
DF_Plot2$Material <- factor(DF_Plot2$Material, levels = c("Plastic", "Glass"), ordered = TRUE)


Tr <- ggplot(DF_Plot2, aes(x = Material, y = Cells_per_g_feces_FR))

Tr <- Tr +
        geom_boxplot(aes(fill = Thaw), outlier.color = NA) +
        geom_point(aes(fill = Thaw, col = Aliquot, shape = Experimentor), position = position_jitterdodge(jitter.height = 0, jitter.width = 0.05), size = 3, alpha = 0.9) +
        theme_bw() +
        scale_color_manual("", values = aliquot_colors) +
        scale_fill_manual("", values = thaw_colors) +
        scale_shape_manual("", values = experimentor_shapes) +
        scale_y_continuous(limits = c(0, max(DF_Plot$Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces] (Volume from set flow rate)") 

Tr <- Tr + stat_compare_means(method = "t.test", hide.ns = FALSE)


# pdf(file = "Person1_t3_FT_Aliquot2_PlasticvsGlass2.pdf", width = 5, height = 4.5)
# Tr
# dev.off()
# --
```


```{r, echo = FALSE, message = FALSE, fig.width = 8, figh.height = 7, warning = FALSE}
Tr
```



## Calculate Mean Counts and so on


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - calculate the mean counts for each Timepoint Thaw, and Aliquot -
DF_Plot <- group_by(DF_Plot, Person, Timepoint, Thaw, Aliquot) 

DF_Plot_summary <- summarise(DF_Plot, Mean_Cells_per_g_feces = mean(Cells_per_g_feces), 
                             SD_Cells_per_g_feces = sd(Cells_per_g_feces),
                             Mean_Cells_per_g_feces_FR = mean(Cells_per_g_feces_FR),
                             SD_Cells_per_g_feces_FR = sd(Cells_per_g_feces_FR))

# -- add further information back --
DF_keep <- unique(dplyr::select(DF_Plot,  Person, Timepoint, Thaw, Aliquot, ExperimentDate:Date_Time))
DF_Plot_summary <- merge(DF_Plot_summary, DF_keep)
# ----
# --


# - calculate the mean counts for each Timepoint and Thaw (ignoring the SD of the aliquots) -
DF_Plot_summary_summary <- group_by(DF_Plot_summary, Person, Timepoint, Thaw)
DF_Plot_summary_summary <- summarise(DF_Plot_summary_summary, Mean_Mean_Cells_per_g_feces = mean(Mean_Cells_per_g_feces), 
                                     SD_Cells_per_g_feces = sd(Mean_Cells_per_g_feces),
                                     SE_Cells_per_g_feces = SD_Cells_per_g_feces/sqrt(n()),
                                     Mean_Mean_Cells_per_g_feces_FR = mean(Mean_Cells_per_g_feces_FR),
                                     SD_Cells_per_g_feces_FR = sd(Mean_Cells_per_g_feces_FR),
                                     SE_Cells_per_g_feces_FR = SD_Cells_per_g_feces_FR/sqrt(n()))

DF_keep <- unique(dplyr::select(DF_Plot_summary,  Person, Timepoint, Thaw, ExperimentDate:Date_Time))
DF_Plot_summary_summary <- merge(DF_Plot_summary_summary, DF_keep)
# --



Tr <- ggplot(DF_Plot_summary_summary, aes(x = Date_Time, y = Mean_Mean_Cells_per_g_feces_FR))

Tr <- Tr +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR-SE_Cells_per_g_feces_FR, ymax = Mean_Mean_Cells_per_g_feces_FR+SE_Cells_per_g_feces_FR, col = Thaw), width = 30000) +
        geom_point(aes(col = Thaw), size = 3, alpha = 0.9) +
        theme_bw() +
        scale_color_manual("", values = thaw_colors) +
        scale_y_continuous(limits = c(0, max(DF_Plot_summary_summary$Mean_Mean_Cells_per_g_feces_FR + DF_Plot_summary_summary$SE_Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces] (Volume from set flow rate)") 
```


```{r, echo = FALSE, message = FALSE, fig.width = 8, figh.height = 7, warning = FALSE}
# Tr
```


## Isolate the Controls and save all data frames from this Person


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# - save the control data to compare between the runs -
DF_control <- dplyr::filter(dplyr::select(DF, Timepoint:`EXPORT TIME`, RecTime_sec:ExperimentDate), Person == "Control")
DF_control$FromPerson <- Person
# --


# - Save the data -
savepath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/PILOT_AfterFlow_Analysis_FInalCounts/Input_Data/Person_Count_rds/"
resList <- list(DF_Plot = DF_Plot, DF_Plot_summary = DF_Plot_summary, DF_Plot_summary_summary = DF_Plot_summary_summary, DF_control = DF_control)

savename <- paste0(Experiment_Date, "_CountAnalysis_", Person, ".rds")

saveRDS(resList, file = file.path(savepath, savename))
# --
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
rm(list = ls())
```





