# Combine Data of all 5 persons


## Load functions and set colors
<!-- all lines where user user input is required or possible are marked with NB: USER INPUT and NB: user input, respectively -->


```{r, echo = FALSE, message=FALSE}
# - source all functions -
functionpath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/Functions_FC/"
function_file <- "20190103_FlowCytometry_Functions.R"
source(file.path(functionpath, function_file))
# --

# - for ggpubr -
symnum.args = list(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))
# -- 

# - Set Colors -
aliquot_colors <- QuantColors15[1:3]
names(aliquot_colors) <- c("Aliquot 1", "Aliquot 2", "Aliquot 3")

experimentor_shapes <- c(16, 17)
names(experimentor_shapes) <- c("Liwei", "Sandra")

# person_colors <- c(cbPalette[2:4], cbPalette[7:8])
# names(person_colors) <- levels(DF$Person)
thaw_colors <- c("#d3d3d3", cbPalette[1])
names(thaw_colors) <- c("FT", "ST")


person_colors <- QuantColors15[1:5]
names(person_colors) <- c("P1", "P2", "P3", "P4", "P5")


timepoint_shapes <- c(16, 17, 15, 18)
names(timepoint_shapes) <- c("t1", "t2", "t3", "t4")
# --
```



## Load the data in

```{r, echo = FALSE, message=FALSE, warning=FALSE}
datapath <- "/Users/jvb740/MarieCurie_Work/CellCounting_TaskForce/FlowCytometryAnalysis/PILOT_Study/PILOT_AfterFlow_Analysis_FInalCounts/Input_Data/Person_Count_rds/"
files <- list.files(datapath)
files <- files[grepl(pattern = "CountAnalysis", files)]

resDFs <- list()

for (i in 1:length(files)){
        resList <- readRDS(file.path(datapath, files[i]))
        resDFs[[i]] <- resList[["DF_Plot_summary_summary"]]
}

# -- restrict to shared columns since Person 5 was made with a different protocol: USE Reduce!! --
colnamesList <- lapply(resDFs, colnames)
keepColnames <- Reduce(intersect, colnamesList)
# ----

resDFs <- lapply(resDFs, dplyr::select, keepColnames)

DF <- do.call("rbind", resDFs)

DF$Person <- factor(DF$Person, levels = c("P1", "P2", "P3", "P4", "P5"), ordered = T)
```



## Plot the mean counts +- standard error vs the collection dates



```{r, echo = FALSE, message = FALSE, warning = FALSE}
Tr <- ggplot(DF, aes(x = Date_Time, y = Mean_Mean_Cells_per_g_feces_FR))

Tr <- Tr +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR-SE_Cells_per_g_feces_FR, ymax = Mean_Mean_Cells_per_g_feces_FR+SE_Cells_per_g_feces_FR, col = Person), width = 30000) +
        # geom_line(aes(col = Person, group = Thaw)) +
        geom_point(aes(col = Person, shape = Thaw), size = 3, alpha = 0.9) +
        theme_bw() +
        facet_grid(Person ~ .) +
        scale_color_manual("", values = person_colors) +
        scale_y_continuous(limits = c(0, max(DF$Mean_Mean_Cells_per_g_feces_FR + DF$SE_Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces]")


# pdf("Combined_Persons_FTvsST.pdf", width = 7, height = 5.5)
# Tr
# dev.off()



Tr1 <- ggplot(DF, aes(x = Date_Time, y = Mean_Mean_Cells_per_g_feces_FR))

Tr1 <- Tr1 +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR-SE_Cells_per_g_feces_FR, ymax = Mean_Mean_Cells_per_g_feces_FR+SE_Cells_per_g_feces_FR, col = Person), width = 30000) +
        geom_line(aes(col = Person)) +
        geom_point(aes(col = Person, shape = Thaw), size = 3, alpha = 0.9) +
        theme_bw() +
        facet_grid(Thaw ~ .) +
        scale_color_manual("", values = person_colors) +
        scale_y_continuous(limits = c(0, max(DF$Mean_Mean_Cells_per_g_feces_FR + DF$SE_Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces]")

# pdf("Combined_Persons_PersonvsPersons.pdf", width = 7, height = 5.5)
# Tr1
# dev.off()

```



```{r, echo = FALSE, message = FALSE, fig.width = 12, figh.height = 14, warning = FALSE}
Tr
```


```{r, echo = FALSE, message = FALSE, fig.width = 12, figh.height = 14, warning = FALSE}
Tr1
```


- **Conclusion: Overall I think it is fair to say that we saw less variation over the 2 weeks than what was reported for many participants by Vandeputte et al. I think your minor changes are more realistic, and we needed to take moisture content into consideration for the weighing process**



## Plot the mean counts +- standard error vs the timepoints



```{r, echo = FALSE, message = FALSE, warning = FALSE}
Tr2 <- ggplot(DF, aes(x = Timepoint, y = Mean_Mean_Cells_per_g_feces_FR))

Tr2 <- Tr2 +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR-SE_Cells_per_g_feces_FR, ymax = Mean_Mean_Cells_per_g_feces_FR+SE_Cells_per_g_feces_FR, col = Person), width = .2) +
        # geom_line(aes(col = Person, group = Thaw)) +
        geom_point(aes(col = Person, shape = Thaw), size = 3, alpha = 0.9) +
        theme_bw() +
        facet_grid(Person ~ .) +
        scale_color_manual("", values = person_colors) +
        scale_y_continuous(limits = c(0, max(DF$Mean_Mean_Cells_per_g_feces_FR + DF$SE_Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces]")



# pdf("Combined_Persons_FTvsST_timepoints.pdf", width = 7, height = 5.5)
# Tr2
# dev.off()



Tr3 <- ggplot(DF, aes(x = Timepoint, y = Mean_Mean_Cells_per_g_feces_FR))

Tr3 <- Tr3 +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR-SE_Cells_per_g_feces_FR, ymax = Mean_Mean_Cells_per_g_feces_FR+SE_Cells_per_g_feces_FR, col = Person), width = .2) +
        geom_line(aes(col = Person, group = Person)) +
        geom_point(aes(col = Person, shape = Thaw), size = 3, alpha = 0.9) +
        theme_bw() +
        facet_grid(Thaw ~ .) +
        scale_color_manual("", values = person_colors) +
        scale_y_continuous(limits = c(0, max(DF$Mean_Mean_Cells_per_g_feces_FR + DF$SE_Cells_per_g_feces_FR))) +
        xlab("") +
        ylab("Cells [per g feces]")
```


```{r, echo = FALSE, message = FALSE, fig.width = 12, figh.height = 14, warning = FALSE}
Tr2
```


```{r, echo = FALSE, message = FALSE, fig.width = 12, figh.height = 14, warning = FALSE}
Tr3
```




## Check correlation between first and second thaw



```{r, echo = FALSE, message = FALSE, warning = FALSE}
DF_Cor <- dplyr::select(DF, Person:Thaw, Mean_Mean_Cells_per_g_feces_FR, SE_Cells_per_g_feces_FR)

# For the next command see:
# <https://stackoverflow.com/questions/29775461/how-can-i-spread-repeated-measures-of-multiple-variables-into-wide-format> 

DF_Cor_w <- recast(DF_Cor, Person + Timepoint ~ variable + Thaw)

UnweightedPearson_Cor <-cor(x = DF_Cor_w$Mean_Mean_Cells_per_g_feces_FR_FT, y = DF_Cor_w$Mean_Mean_Cells_per_g_feces_FR_ST, method = "pearson")

PVal_Perason = cor.test(x = DF_Cor_w$Mean_Mean_Cells_per_g_feces_FR_FT, y = DF_Cor_w$Mean_Mean_Cells_per_g_feces_FR_ST, method = "pearson")



Tr <- ggplot(DF_Cor_w, aes(x = Mean_Mean_Cells_per_g_feces_FR_FT, y = Mean_Mean_Cells_per_g_feces_FR_ST))

Tr <- Tr +
        geom_errorbar(aes(ymin = Mean_Mean_Cells_per_g_feces_FR_ST-SE_Cells_per_g_feces_FR_ST, ymax = Mean_Mean_Cells_per_g_feces_FR_ST+SE_Cells_per_g_feces_FR_ST, col = Person)) +
        geom_errorbarh(aes(xmin = Mean_Mean_Cells_per_g_feces_FR_FT-SE_Cells_per_g_feces_FR_FT, xmax = Mean_Mean_Cells_per_g_feces_FR_FT+SE_Cells_per_g_feces_FR_FT, col = Person)) +
        # geom_line(aes(col = Person, group = Thaw)) +
        geom_point(aes(col = Person, shape = Timepoint), size = 3, alpha = 0.9) +
        geom_abline(slope = 1,intercept = 0, col = "black", lty = 2) +
        geom_smooth(method = "lm", lty = 4) +
        theme_bw() +
        scale_color_manual("", values = person_colors) +
        scale_shape_manual("", values = timepoint_shapes) +
        scale_y_continuous(limits = c(0, 1.85e10)) +
        scale_x_continuous(limits = c(0, 1.85e10)) +
        #scale_x_log10(limits = c(1e9, 2e10)) +
        #scale_y_log10(limits = c(1e9, 2e10)) +
        xlab("Cells [per g feces] First Thaw") +
        ylab("Cells [per g feces] Second Thaw") +
        coord_equal(ratio = 1) +
        ggtitle(paste0("Pearson Correlation: ", round(UnweightedPearson_Cor,3)))



# pdf("Correlation_FT_ST.pdf", width = 6, height = 6)
# Tr
# dev.off()

```


```{r, echo = FALSE, message = FALSE, fig.width = 8, figh.height = 8, warning = FALSE}
Tr
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
rm(list = ls())
```










